---
title: "Vignette for Meta Sparse Kmeans"
author: "Zhiguang Huo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  \VignetteIndexEntry{MetaSparseKmeans}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`MetaSparseKmeans` aims to perform sample clustering on multiple transcriptomic studies.
Three goals are acheived simultanously.

1. Sample clustering will be performed in each study.
2. Clustering pattern will be consistent across studies.
3. A small subset of intrinsic genes with the will be selected.

## Reference
Huo, Z., Ding, Y., Liu, S., Oesterreich, S., & Tseng, G. (2016). Meta-analytic framework for sparse K-means to identify disease subtypes in multiple transcriptomic studies. Journal of the American Statistical Association, 111(513), 27-42.

## How to install this R package?
Refer to [https://github.com/Caleb-Huo/MetaSparseKmeans](https://github.com/Caleb-Huo/MetaSparseKmeans)

## Important input

- x: A list for several microarray studies. Each element of the list should be a p*n matrix. p is number of features and n is number of samples. Clustering is performed on sample level. p has to be the same for each study. Missing value should be set to be NA. Current version won't support missing value, it will be allowed in the next version.

- K: K specifies number of clusters. We assume the number of clusters to be the same in each study.

- wbounds: wbounds is the tuning parameter that controls number of selected features. Larger tuning parameter yield more selected features. wbounds could be a number or a vector. wbounds is suggested to be selected using prior information (e.g. which tuning parameter generate best survival difference.)

## Tuning parameter:

- K: suggest to be selected in each individual study (after filtering) using gap statistics. Then choose a common K.
- wbounds: can be tuned using `gapStat_MSKM()` function. However, in real data, this approach is time comsuming and often ends up with more genes. Users are suggested to use a sequence of wbounds and select the best one with the best biological interpretation.
- lambda: suggest to use the default value 1/2.

## Usage

### generate two studies.

```{r}
library(MetaSparseKmeans)
```

```{r}
set.seed(15213)

G = 1000
n11 = 100
n12 = 100
n13 = 150
label1 = c(rep(1,n11),rep(2,n12),rep(3,n13))

P0 = 0.6
P1 = 0.1
P2 = 0.1
P3 = 0.1
P4 = 0.1
sd = 0.5

G0 = G*P0  # nonDE genes
G1 = G*P1  # DE H-L
G2 = G*P2  # DE L-H
G3 = G*P3
G4 = G*P4


mu111 = runif(G1,-0.25,0.25)
mu112 = runif(G1,0.5,1)
mu113 = runif(G1,-1,-0.5)

mu121 = runif(G2,-1,-0.5)
mu122 = runif(G2,-0.25,0.25)
mu123 = runif(G2,0.5,1)

mu131 = runif(G3,-1,-0.5)
mu132 = runif(G3,-0.25,0.25)
mu133 = runif(G3,0.5,1)

mu14 = runif(G4,-0.25,0.25)
mu10 = runif(G0,-0.25,0.25)

Data111 = matrix(rnorm(n11*G1,mu111,sd^2),nrow=G1)
Data112 = matrix(rnorm(n12*G1,mu112,sd^2),nrow=G1)
Data113 = matrix(rnorm(n13*G1,mu113,sd^2),nrow=G1)
Data11 = cbind(Data111,Data112,Data113)

Data121 = matrix(rnorm(n11*G2,mu121,sd^2),nrow=G2)
Data122 = matrix(rnorm(n12*G2,mu122,sd^2),nrow=G2)
Data123 = matrix(rnorm(n13*G2,mu123,sd^2),nrow=G2)
Data12 = cbind(Data121,Data122,Data123)

Data131 = matrix(rnorm(n11*G3,mu131,sd^2),nrow=G3)
Data132 = matrix(rnorm(n12*G3,mu132,sd^2),nrow=G3)
Data133 = matrix(rnorm(n13*G3,mu133,sd^2),nrow=G3)
Data13 = cbind(Data131,Data132,Data133)

Data14 = matrix(rnorm((n11+n12+n13)*G4,mu14,sd^2),nrow=G4)

Data10 = matrix(rnorm((n11+n12+n13)*G0,mu10,sd^2),nrow=G0)

S1 = rbind(Data10,Data11,Data12,Data13,Data14)


G = 1000
n21 = 150
n22 = 100
n23 = 100

label2 = c(rep(1,n21),rep(2,n22),rep(3,n23))

P0 = 0.6
P1 = 0.1 #common features
P2 = 0.1 #common features
P3 = 0.1 #noise in S1
P4 = 0.1 #noise in S2
sd = 0.5

G0 = G*P0  # nonDE genes
G1 = G*P1  # DE H-L
G2 = G*P2  # DE L-H
G3 = G*P3  #noise in S1
G4 = G*P4  #noise in S2

mu211 = runif(G1,-0.25,0.25)
mu212 = runif(G1,0.5,1)
mu213 = runif(G1,-1,-0.5)

mu221 = runif(G2,-1,-0.5)
mu222 = runif(G2,-0.25,0.25)
mu223 = runif(G2,0.5,1)

mu23 = runif(G3,-0.25,0.25)

mu241 = runif(G4,-1,-0.5)
mu242 = runif(G4,-0.25,0.25)
mu243 = runif(G4,0.5,1)

mu20 = runif(G0,-0.25,0.25)

Data211 = matrix(rnorm(n21*G1,mu211,sd^2),nrow=G1)
Data212 = matrix(rnorm(n22*G1,mu212,sd^2),nrow=G1)
Data213 = matrix(rnorm(n23*G1,mu213,sd^2),nrow=G1)
Data21 = cbind(Data211,Data212,Data213)

Data221 = matrix(rnorm(n21*G2,mu221,sd^2),nrow=G2)
Data222 = matrix(rnorm(n22*G2,mu222,sd^2),nrow=G2)
Data223 = matrix(rnorm(n23*G2,mu223,sd^2),nrow=G2)
Data22 = cbind(Data221,Data222,Data223)

Data23 = matrix(rnorm((n21+n22+n23)*G3,mu23,sd^2),nrow=G3)

Data241 = matrix(rnorm(n21*G4,mu241,sd^2),nrow=G4)
Data242 = matrix(rnorm(n22*G4,mu242,sd^2),nrow=G4)
Data243 = matrix(rnorm(n23*G4,mu243,sd^2),nrow=G4)
Data24 = cbind(Data241,Data242,Data243)


Data20 = matrix(rnorm((n21+n22+n23)*G0,mu20,sd^2),nrow=G0)

S2 = rbind(Data20,Data21,Data22,Data23,Data24)
S = list(t(S1),t(S2))
```

### visualize the data

```{r}
getWsHeatmap(t(S[[1]]),label1)

getWsHeatmap(t(S[[2]]),label2)

```

### perform meta sparse Kmeans
```{r}

res = MetaSparseKmeans(x=S,K=3,wbounds=10,lambda=0.5)
```

### visualize the result

```{r}

getWsHeatmap(t(S[[1]]),res$Cs[[1]],res$ws)

getWsHeatmap(t(S[[2]]),res$Cs[[2]],res$ws)

plot(res$ws,xlab='geneIndex')
```

